Certified Symbolic Finite Transducers: Formalization and Applications to String Analysis

### 2025-12-01 Paper/application.tex
camera-ready

Listing \ref{lst-html-encode} presents a C\# code snippet implementing an HTML encoding function for Anti-XSS. This function iterates through each character in the input string and encodes characters that are unsafe for HTML rendering.

\begin{lstlisting}[language={[Sharp]C}, caption={C\# Code for AntiXSS.EncodeHtml version 2.0.}, label={lst-html-encode}, float=htbp]
static string EncodeHtml(string t)
{
  if (t == null) { return null; }
  if (t.Length == 0) {
    return string.Empty;
  }
  StringBuilder builder = new
    StringBuilder("", t.Length * 2);
  foreach (char c in t) {
    if ((((c > ''') && (c < '{')) |$\mid\mid$|
         ((c > '@') && (c < '['))) |$\mid\mid$|
         (((c == ' ') || ((c > '/') && (c < ':'))) |$\mid\mid$|
         (((c == '.') || (c == ',')) |$\mid\mid$|
         ((c == '-') || (c == '_'))))) {
          builder.Append(c);
      }
      else {
          builder.Append("&#" +
          ((int) c).ToString() + ";");
      }
  }
  return builder.ToString();
}
\end{lstlisting}

@@

\begin{table}[h]
  \centering
  \small
  \setlength{\tabcolsep}{4pt} % Reduced column gap
  \begin{tabular}{p{2.2cm}ll} % Increased width for automatic line break
    \toprule
    \textbf{Attack Model} & \textbf{Verfication Result} & \textbf{Size} \\
    \midrule
    attack\_html           & GA HtmlEscape: \textbf{Unsafe} & 1/6\\
                           & Trim + GA HtmlEscape: \textbf{Safe} & 4/12 \\
                           & escapeString: \textbf{Unsafe} & 1/11 \\
                           & Trim + escapeString: \textbf{Safe} & 4/17 \\
                           & Trim + OWASP HTMLEncode: \textbf{Safe} & 4/18 \\
    \midrule
    attack\_javascript     & GA HtmlEscape + GA PreEscape + & 3/20 \\
                           &  gaSnippetesc: \textbf{safe} & \\
    \midrule
    attack\_json           & Trim + GA JSONESc : \textbf{Safe} & 4/16\\
    \midrule
    attack\_xml            & Trim + GA XMLEsc: \textbf{safe} & 4/16\\
    \midrule
    attack\_ccs            & Trim + GA CleanseCSS : \textbf{Safe} & 25/46\\
    \bottomrule
  \end{tabular}
  \caption{Sanitizer verfication result}
  \label{tab:sanitizer_attack}
\end{table}

@@

\begin{lstlisting}[language={}, caption={CSS attack model for sanitizer verification.}, label={lst-css-attack}, float=htbp]
|$\Sigma$|* ( / \* [\s\S]*? \*/
  |$\mid$| ; |$\mid$| \{ |$\mid$| \}
  |$\mid$| [eE][xX][pP][rR][eE][sS][sS][iI]
    [oO][nN]\s*\(
  |$\mid$| @\s*[iI][mM][pP][oO][rR][tT]
  |$\mid$| [bB][eE][hH][aA][vV][iI][oO][rR]\s*:
  |$\mid$| -\s*[mM][oO][zZ]-\s*[bB][iI][nN]
    [dD][iI][nN][gG]
  |$\mid$| (?!allowUrl) [uU][rR][lL]\s*\(
 ) |$\Sigma$|*
\end{lstlisting}

@@

\begin{table*}[h]
\centering
\small
\setlength{\tabcolsep}{4pt}
\renewcommand{\arraystretch}{1.1}
\begin{tabular}{llccccccc}
  \toprule
  \textbf{Subgroup} & \textbf{Test} & \textbf{Count} & \textbf{SAT} & \textbf{UNSAT} & \textbf{Inconclusive} & \textbf{CertiStrR (s)} & \textbf{Ostrich (s/timeout)} & \textbf{CVC5 (s/timeout)} \\
  \midrule
  \multirow{2}{*}{Replace-First} & Replace RE  & 94  & 77  & 16  & 1 & 0.79 & 166.27/27 & 2.67/20 \\
   & Replace Str & 323 & 103 & 211 & 9 & 1.94 & 7.10/20 & 0.81/1 \\
     \midrule
  \multirow{4}{*}{Replace-All} & PCP & 1000 & 751 & 249 & 0 & 0.14 & 3.412/0 & unknown \\
   & RNA SAT   & 500  & 500 & 0   & 0 & 2.81 & 55.61/0 & unknown \\
   & RNA UNSAT & 500  & 0   & 500 & 0 & 3.14 & 64.185/0 & unknown \\
   & WebApp      & 255  & 40  & 215 & 0 & 0.79 & 23.67/17 & 8.379/1 \\
  \bottomrule
\end{tabular}
\caption{Experimental results}
\label{tab:string_operations}
\end{table*}

@@

\begin{table}[h]
  \centering
  \small
  \begin{tabular}{lccc}
      \toprule
      & \textbf{Definitions} & \textbf{Lemmas} & \textbf{Proofs (LOC)} \\
      \midrule
      Abstract & 17 & 21 & 3274 \\
      Implementation & 51 & 43 & 2700 \\
      Interval & 15 & 29 & 1500 \\
      \bottomrule
  \end{tabular}
  \caption{Overview of the effort of certified development}
  \label{tab:abstract_impl}
\end{table}

>>>>

Figure \ref{lst-html-encode} presents a C\# code snippet implementing an HTML encoding function for Anti-XSS. This function iterates through each character in the input string and encodes characters that are unsafe for HTML rendering.

\begin{figure}[hbt!]
\begin{lstlisting}[language={[Sharp]C}]
static string EncodeHtml(string t)
{
  if (t == null) { return null; }
  if (t.Length == 0) {
    return string.Empty;
  }
  StringBuilder builder = new
    StringBuilder("", t.Length * 2);
  foreach (char c in t) {
    if ((((c > ''') && (c < '{')) |$\mid\mid$|
         ((c > '@') && (c < '['))) |$\mid\mid$|
         (((c == ' ') || ((c > '/') && (c < ':'))) |$\mid\mid$|
         (((c == '.') || (c == ',')) |$\mid\mid$|
         ((c == '-') || (c == '_'))))) {
          builder.Append(c);
      }
      else {
          builder.Append("&#" +
          ((int) c).ToString() + ";");
      }
  }
  return builder.ToString();
}
\end{lstlisting}
\caption{C\# Code for AntiXSS.EncodeHtml version 2.0.}
\label{lst-html-encode}
\end{figure}

@@

\begin{table}[h]
  \centering
  \small
  \caption{Sanitizer verfication result}
  \label{tab:sanitizer_attack}
  \setlength{\tabcolsep}{4pt} % Reduced column gap
  \begin{tabular}{p{2.2cm}ll} % Increased width for automatic line break
    \toprule
    \textbf{Attack Model} & \textbf{Verfication Result} & \textbf{Size} \\
    \midrule
    attack\_html           & GA HtmlEscape: \textbf{Unsafe} & 1/6\\
                           & Trim + GA HtmlEscape: \textbf{Safe} & 4/12 \\
                           & escapeString: \textbf{Unsafe} & 1/11 \\
                           & Trim + escapeString: \textbf{Safe} & 4/17 \\
                           & Trim + OWASP HTMLEncode: \textbf{Safe} & 4/18 \\
    \midrule
    attack\_javascript     & GA HtmlEscape + GA PreEscape + & 3/20 \\
                           &  gaSnippetesc: \textbf{safe} & \\
    \midrule
    attack\_json           & Trim + GA JSONESc : \textbf{Safe} & 4/16\\
    \midrule
    attack\_xml            & Trim + GA XMLEsc: \textbf{safe} & 4/16\\
    \midrule
    attack\_ccs            & Trim + GA CleanseCSS : \textbf{Safe} & 25/46\\
    \bottomrule
  \end{tabular}
\end{table}

@@

\begin{lstlisting}[language={}, caption={CSS attack model for sanitizer verification}, captionpos=b, label={lst-css-attack}, float=htbp]
|$\Sigma$|* ( / \* [\s\S]*? \*/
  |$\mid$| ; |$\mid$| \{ |$\mid$| \}
  |$\mid$| [eE][xX][pP][rR][eE][sS][sS][iI]
    [oO][nN]\s*\(
  |$\mid$| @\s*[iI][mM][pP][oO][rR][tT]
  |$\mid$| [bB][eE][hH][aA][vV][iI][oO][rR]\s*:
  |$\mid$| -\s*[mM][oO][zZ]-\s*[bB][iI][nN]
    [dD][iI][nN][gG]
  |$\mid$| (?!allowUrl) [uU][rR][lL]\s*\(
 ) |$\Sigma$|*
\end{lstlisting}

@@

\begin{table*}[h]
\centering
\small
\caption{Experimental results}
\label{tab:string_operations}
\setlength{\tabcolsep}{4pt}
\renewcommand{\arraystretch}{1.1}
\begin{tabular}{llccccccc}
  \toprule
  \textbf{Subgroup} & \textbf{Test} & \textbf{Count} & \textbf{SAT} & \textbf{UNSAT} & \textbf{Inconclusive} & \textbf{CertiStrR (s)} & \textbf{Ostrich (s/timeout)} & \textbf{CVC5 (s/timeout)} \\
  \midrule
  \multirow{2}{*}{Replace-First} & Replace RE  & 94  & 77  & 16  & 1 & 0.79 & 166.27/27 & 2.67/20 \\
   & Replace Str & 323 & 103 & 211 & 9 & 1.94 & 7.10/20 & 0.81/1 \\
     \midrule
  \multirow{4}{*}{Replace-All} & PCP & 1000 & 751 & 249 & 0 & 0.14 & 3.412/0 & unknown \\
   & RNA SAT   & 500  & 500 & 0   & 0 & 2.81 & 55.61/0 & unknown \\
   & RNA UNSAT & 500  & 0   & 500 & 0 & 3.14 & 64.185/0 & unknown \\
   & WebApp      & 255  & 40  & 215 & 0 & 0.79 & 23.67/17 & 8.379/1 \\
  \bottomrule
\end{tabular}
\end{table*}

@@

\begin{table}[h]
  \centering
  \small
  \caption{Overview of the effort of certified development}
  \label{tab:abstract_impl}
  \begin{tabular}{lccc}
      \toprule
      & \textbf{Definitions} & \textbf{Lemmas} & \textbf{Proofs (LOC)} \\
      \midrule
      Abstract & 17 & 21 & 3274 \\
      Implementation & 51 & 43 & 2700 \\
      Interval & 15 & 29 & 1500 \\
      \bottomrule
  \end{tabular}
\end{table}

### 2025-12-01 Paper/formalization.tex
camera-ready

\paragraph{Implementation layer of the product operation.} We now present the algorithmic implementation of the product operation between an SFT and an SFA\footnote{For clarity of presentation, we show a simplified version of the Isabelle/HOL implementation while preserving the essential algorithmic structure.} based on Refined\_monadic framework.

@@

  The correctness of the product computation is established through a refinement proof, demonstrating that \texttt{productT\_imp} (Figure \ref{fig-compute-nft-product}) correctly implements the abstract specification \texttt{productT} (Figure \ref{fig-def-FTProd}). This refinement relationship is formally specified in Figure \ref{fig-def-productT_imp_correct}, where we leverage the \emph{Refine\_Monadic} framework's data refinement.

>>>>

\paragraph{Implementation layer of the product operation.} We now present the algorithmic implementation of the product operation between an SFT and an SFA based on Refined\_monadic framework.

@@

  The correctness of the product operation is established by a refinement proof, demonstrating that \texttt{productT\_imp} (Figure \ref{fig-compute-nft-product}) correctly implements the abstract specification \texttt{productT} (Figure \ref{fig-def-FTProd}). This refinement relationship is formally specified in Figure \ref{fig-def-productT_imp_correct}, where we leverage the \emph{Refine\_Monadic} framework's data refinement.

### 2025-12-01 Paper/main.tex
fix reviews
camera-ready

\documentclass[sigplan,10pt]{acmart}\settopmatter{printfolios=true,printccs=false,printacmref=false}
%This is a template for producing LIPIcs articles. 
%See lipics-v2021-authors-guidelines.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
%for section-numbered lemmas etc., use "numberwithinsect"
%for enabling cleveref support, use "cleveref"
%for enabling autoref support, use "autoref"
%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
%for enabling thm-restate support, use "thm-restate"
%for enabling a two-column layout for the author/affilation part (only applicable for > 6 authors), use "authorcolumns"
%for producing a PDF according the PDF/A standard, add "pdfa"

%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository
\usepackage{stmaryrd}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usetikzlibrary{intersections}
\usetikzlibrary{automata}
\usetikzlibrary{positioning}
\usepackage{booktabs}
%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory
\usepackage{listings}
\usepackage{xcolor}
\usepackage{multirow}

@@

\lstdefinelanguage{SMTLIB}{
  morekeywords={set-logic, declare-fun, assert, check-sat, get-model},
  sensitive=true,
  morecomment=[l];,
  morestring=[b]"
}
\lstset{
  language=SMTLIB,
  basicstyle=\ttfamily\small,
  keywordstyle=\color{blue}\bfseries,
  commentstyle=\color{green!50!black},
  numbers=left,
  numberstyle=\tiny\color{gray},
  stepnumber=1,
  numbersep=5pt,
  backgroundcolor=\color{gray!10},
  frame=single,
  breaklines=true,
  captionpos=b,
  showstringspaces=false
}

\bibliographystyle{plainurl}% the mandatory bibstyle

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\lstdefinelanguage{Isabelle}{
    keywords={definition,theorem,lemma,proof,qed,by,auto,simp,apply,done,
              where,fixes,assumes,shows,if,then,else,let,in,case,of,
              record,datatype,type_synonym,fun,function,primrec},
    sensitive=true,
    morecomment=[s]{(*}{*)},
    morestring=[b]",
  }

\lstset{
  language=Isabelle,
	basicstyle=\ttfamily,
	keywordstyle=\color{blue}\bfseries,
	morekeywords={definition,if,then,else,where,record,fun,lemma,type_synonym,fixes,shows,assumes,and, locale},
	escapeinside={|}{|},
  sensitive=true
}

@@

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}
\author{Shuanglong Kan}
\affiliation{
  \institution{Barkhausen Institute}
  \country{Germany}
}

\author{Anthony W. Lin}
\affiliation{
  \institution{TU Kaiserslautern \& MPI-SWS}
  \country{Germany}
}

\title{Certified Symbolic Finite Transducers: Formalization and Applications to
String Analysis}

\input{abstract.tex}
\maketitle

@@

\bibliography{literature}
\end{document}

>>>>

\documentclass[sigplan,screen]{acmart}
%This is a template for producing LIPIcs articles. 
%See lipics-v2021-authors-guidelines.pdf for further information.
%for A4 paper format use option "a4paper", for US-letter use option "letterpaper"
%for british hyphenation rules use option "UKenglish", for american hyphenation rules use option "USenglish"
%for section-numbered lemmas etc., use "numberwithinsect"
%for enabling cleveref support, use "cleveref"
%for enabling autoref support, use "autoref"
%for anonymousing the authors (e.g. for double-blind review), add "anonymous"
%for enabling thm-restate support, use "thm-restate"
%for enabling a two-column layout for the author/affilation part (only applicable for > 6 authors), use "authorcolumns"
%for producing a PDF according the PDF/A standard, add "pdfa"

%\pdfoutput=1 %uncomment to ensure pdflatex processing (mandatatory e.g. to submit to arXiv)
%\hideLIPIcs  %uncomment to remove references to LIPIcs series (logo, DOI, ...), e.g. when preparing a pre-final version to be uploaded to arXiv or another public repository
\usepackage{stmaryrd}
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usetikzlibrary{intersections}
\usetikzlibrary{automata}
\usetikzlibrary{positioning}
\usepackage{booktabs}
%\graphicspath{{./graphics/}}%helpful if your graphic files are in another directory
\usepackage{listings}
\usepackage{xcolor}
\usepackage{multirow}
\usepackage{balance}
\usepackage{microtype}

@@

\lstdefinelanguage{SMTLIB}{
  morekeywords={set-logic, declare-fun, assert, check-sat, get-model},
  sensitive=true,
  morecomment=[l];,
  morestring=[b]"
}

\lstdefinelanguage{Isabelle}{
    keywords={definition,theorem,lemma,proof,qed,by,auto,simp,apply,done,
              where,fixes,assumes,shows,if,then,else,let,in,case,of,
              record,datatype,type_synonym,fun,function,primrec},
    sensitive=true,
    morecomment=[s]{(*}{*)},
    morestring=[b]",
  }

\lstset{
  basicstyle=\ttfamily,
  keywordstyle=\color{blue}\bfseries,
  morekeywords={definition,if,then,else,where,record,fun,lemma,type_synonym,fixes,shows,assumes,and,locale,set-logic,declare-fun,assert,check-sat,get-model},
  escapeinside={|}{|},
  sensitive=true,
  commentstyle=\color{green!50!black},
  morecomment=[l];,
  morestring=[b]",
  breaklines=true,
  showstringspaces=false
}

\bibliographystyle{ACM-Reference-Format}% the mandatory bibstyle

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\title[Certified Symbolic Finite Transducers]{Certified Symbolic Finite Transducers: Formalization and Applications to
String Analysis}

\special{papersize=8.5in,11in}
\setlength{\pdfpageheight}{\paperheight}
\setlength{\pdfpagewidth}{\paperwidth}

\author{Shuanglong Kan}
\orcid{0000-0002-4132-2626}
\affiliation{%
  \department{ Verified System Design Automation Group}
  \institution{Barkhausen Institut}
  \city{Dresden}
  \country{Germany}
}
\email{shuanglong.kan@barkhauseninstitut.org}

\author{Anthony W. Lin}
\orcid{0000-0003-4715-5096}
\affiliation{
  \department{Department of Computer Science}
  \institution{University of Kaiserslautern-Landau \& MPI-SWS}
  \city{Kaiserslautern}
  \country{Germany}
}
\email{awlin@mpi-sws.org}

\input{abstract.tex}

\begin{CCSXML}
<ccs2012>
   <concept>
       <concept_id>10003752.10003790.10003794</concept_id>
       <concept_desc>Theory of computation~Automated reasoning</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
   <concept>
       <concept_id>10003752.10003766.10003773.10003774</concept_id>
       <concept_desc>Theory of computation~Transducers</concept_desc>
       <concept_significance>500</concept_significance>
       </concept>
 </ccs2012>
\end{CCSXML}

\ccsdesc[500]{Theory of computation~Automated reasoning}
\ccsdesc[500]{Theory of computation~Transducers}

\keywords{symbolic transducers, Isabelle/HOL, string theory, SMT solvers} %TODO mandatory; please add comma-separated list of keywords
\maketitle

@@

\balance
\bibliography{literature}
\end{document}
